Operational Feature Selection in Gaussian Mixture Models

### 2016-09-30 paper.tex
update

    To predict the  class of a new unseen sample,  the maximum \emph{a
      posteriori}  rule  is  used:
    \begin{equation*}
        \mathbf{x} \text{ belongs to } c \Leftrightarrow c = \text{arg} \max_{c \in C} p(c) p(\mathbf{x}|c).
    \end{equation*}
    Under the GMM,  and identifying $p(c)$ as $\pi_c$  and $p(x|c)$ as
    $f_c(x|\theta)$ and by taking the log, a decision rule is obtained
    \begin{eqnarray}\label{eq:decision}
      Q_c(\mathbf{x}) &=& 2 \log \left( p(c) p(\mathbf{x}|c) \right) \nonumber \\
                      &=& - (\mathbf{x} - \boldsymbol{\mu}_c)^t \boldsymbol{\Sigma}_c^{-1} (\mathbf{x} - \boldsymbol{\mu}_c) \nonumber \\
                      & &-\log (|\boldsymbol{\Sigma}_c|) + 2 \log (\pi_c) - d \log (2\pi).
    \end{eqnarray}

@@

    %% TODO: il faut reprendre cette partie en détaillant plus
    %% Il y a deux approches: une qui regularise le probleme inverse,
    %% ridge est une solution, mais il y en a d'autres. Voir le
    %% paragraphe qui est consacré à cela dans la thèse d'Anthony ->
    %% elle arrive pas mail. Ensuite, il y a la réduction de dimension
    %% données des refs (voir intro du papier JSTARS) et dans la
    %% réduction  de dimension, il  y a  la sélection de  features. La
    %% conclusion doit s'enchainer avec la partie suivante.
    There are two major  solutions to this problem. First option is to stabilize the inversion of the covariance matrices work. Some methods investigate the use of constraints on the direct problem \cite{reynolds1995robust} \cite{hathaway1985constrained} and some work on the inverse problem by using regularization method \cite{hoerl1970ridge} \cite{jensen2008regression}. Second option is to reduce dimension. Features extraction/selection methods have been developed in order to reduce the dimension. In this  study, the latter option is explored and a feature selection method named sequential forward features selection is presented after a brief overview of the selection algorithms in the next Section. %A finir

@@

        A measure of correct classification is based on an error matrix $M$, or \emph{confusion matrix}~\cite[Chapter 4]{congalton2008assessing}. An element $M_{ij}$ of the matrix represents the number of samples of class $i$ classified as class $j$. The confusion matrix allows the computation of several global and per-class indices related to the classification accuracy~\cite{congalton2008assessing}. Three global criteria were used 
        \begin{itemize}
        \item  \emph{The overall  accuracy} (OA)  is the  rate of  the
          number of samples with the  correct predicted label over the
          total  number of  samples~\cite{congalton2008assessing}. This
          metric is  easy to interpret  but is  biased in the  case of
          unbalanced classes.

@@

      The classification rate is  estimated by a $k$- cross-validation
      ($k$-CV)~\cite{opac-b1127878}.  To compute  the $k$-CV, a subset
      is removed  from $\mathcal{S}$ and  the GMM is learned  with the
      remaining training samples.   A test error is  computed with the
      removed  training  samples  used  as  validation  samples.   The
      process is  iterated $k$ times and  the estimated classification
      rate is computed as the mean  test error over the $k$ subsets of
      $\mathcal{S}$.

@@

        \subsubsection{Similarity between distributions}
        These measures are called divergence function and are defined so that, if $S$ is a space of all probability distribution with same support, it satisfies
        \begin{align*}
            &\forall (p,q) \in S, \text{Div}(p,q) \geq 0, \\
            &\text{Div}(p,q) = 0 \Leftrightarrow p = q.
        \end{align*}
        More specifically, focus is made on two particular divergences: the Kullback–Leibler divergence and the Jeffries–Matusita distance. The advantage of these divergences is that they have an explicit expression in the case of Gaussian models. The simplification allows to get rid of any integration calculus which is a major problem when dealing with high-dimensional data.

>>>>

    To predict the  class of a new unseen sample,  the maximum \emph{a
      posteriori}  rule  is  used:
    \begin{equation*}
        \mathbf{x} \text{ belongs to } c \Leftrightarrow c = \text{arg} \max_{c \in C} p(c) p(\mathbf{x}|c).
    \end{equation*}
    Under the GMM,  and identifying $p(c)$ as $\pi_c$  and $p(\mathbf{x}|c)$ as
    $f_c(\mathbf{x}|\theta)$ and by taking the log, a decision rule is obtained
    \begin{eqnarray}\label{eq:decision}
      Q_c(\mathbf{x}) &=& 2 \log \left( p(c) p(\mathbf{x}|c) \right) \nonumber \\
                      &=& - (\mathbf{x} - \boldsymbol{\mu}_c)^t \boldsymbol{\Sigma}_c^{-1} (\mathbf{x} - \boldsymbol{\mu}_c) \nonumber \\
                      & &-\log (|\boldsymbol{\Sigma}_c|) + 2 \log (\pi_c) - d \log (2\pi).
    \end{eqnarray}

@@

    %% TODO: il faut reprendre cette partie en détaillant plus
    %% Il y a deux approches: une qui regularise le probleme inverse,
    %% ridge est une solution, mais il y en a d'autres. Voir le
    %% paragraphe qui est consacré à cela dans la thèse d'Anthony ->
    %% elle arrive pas mail. Ensuite, il y a la réduction de dimension
    %% données des refs (voir intro du papier JSTARS) et dans la
    %% réduction  de dimension, il  y a  la sélection de  features. La
    %% conclusion doit s'enchainer avec la partie suivante.
    There are two major solutions to this problem. First option is to stabilize the inversion of the covariance matrices. Some methods investigate the use of constraints on the direct problem. Reynolds \emph{et al.} \cite{reynolds1995robust} proposed to use diagonal covariance matrices. It is also possible to force the diagonal element to be higher than a given value by maximizing the GMM likelihood \cite{hathaway1985constrained}. Celeux and Govaert \cite{celeux1995gaussian} suggest to use equality constraints between coefficients of the covariance matrix in a parsimonious cluster-based GMM framework. Others papers propose to work on the inverse problem. A classical method is to use a regularization method as the well-known ridge regularization \cite{hoerl1970ridge}. A ridge regularization aims to stabilize the inversion by replacing the covariance matrix $\boldsymbol{\Sigma}_c$ by $\boldsymbol{\Sigma}_c + \tau \mathbf{I}$ where $\tau$ is a positive parameter and $\mathbf{I}$ the identity matrix. Jensen \emph{et al.} \cite{jensen2008regression} propose a different approach using a sparsity approximation to inverse the covariance matrix.

    Second option is to reduce dimension. Features extraction/selection methods have been developed in order to reduce the dimension with various approaches described in Section~\ref{sec:intro}. In this study, this latter option is explored and a feature selection method named sequential forward features selection is presented after a brief overview of the selection algorithms in the next Section.

@@

        A measure of good classification is based on an error matrix $M$, or \emph{confusion matrix}~\cite[Chapter 4]{congalton2008assessing}. An element $M_{ij}$ of the matrix represents the number of samples of class $i$ classified as class $j$. The confusion matrix allows the computation of several global and per-class indices related to the classification accuracy~\cite{congalton2008assessing}. Three global criteria were used
        \begin{itemize}
        \item  \emph{The overall  accuracy} (OA)  is the  rate of  the
          number of samples with the  correct predicted label over the
          total  number of  samples~\cite{congalton2008assessing}. This
          metric is  easy to interpret  but is  biased in the  case of
          unbalanced classes.

@@

        In order to estimate classification rate, a cross-validation process over training set is used\cite{opac-b1127878}. The training set is divided in k folds, then, the training is done with $(k-1)$ folds and the performances are estimated with the remaining fold.

@@

        \subsubsection{Similarity between distributions}

        \hspace{0pt} \\
        The second type of criterion functions is a measure of similarity between two distributions. These measures are called divergence function and are defined so that, if $S$ is a space of all probability distribution with same support, it satisfies
        \begin{align*}
            &\forall (p,q) \in S, \text{Div}(p,q) \geq 0, \\
            &\text{Div}(p,q) = 0 \Leftrightarrow p = q.
        \end{align*}
        More specifically, focus is made on two particular divergences: the Kullback–Leibler divergence and the Jeffries–Matusita distance. The advantage of these divergences is that they have an explicit expression in the case of Gaussian models. The simplification allows to get rid of any integration calculus which is a major problem when dealing with high-dimensional data.
